---
title: "Larvalfishdata"
output: html_notebook
---
```{r setup}
library(tidyverse)
library(lubridate)
library(ggplot2)
NJ_temp <- read.csv("env_17.csv") #NJ excel file of catch temperature
NC_temp <- read.csv("NC_temp_final.csv") #NC excel file of catch temperature
NJ_popcatchid <- read.csv("nj_popdata.final.csv") #NJ file of catch number and id 
NC_popcatchid <- read.csv("nc_popdata.final.csv") # NC file of catch number and id 
NJfishmeasurements <- read.csv("Njfishdataedit.csv") #NJ measurements
NCfishmesurements <- read.csv("Ncfishdatamesurements.csv") #NC mesurements

```

```{r examining data frames}
summary(NC_popcatchid)
str(NC_popcatchid)
str(NJ_popcatchid)
str(NCfishmesurements)
view(NCfishmesurements) #viewsspreed
```
Reduce NJ temp data
```{r NJ temp data reduce}
head(NJ_temp) #first5rows - ,10 to increase rows
#fix date
NJ_temp$Date <- ymd(NJ_temp$Date) #replacementsamename

#reduce to columns we actually need
NJ_temp.r <- NJ_temp %>%
  select(Date, SuTemp)

#also, add region column
NJ_temp.r$region <- "NJ"
```

Reduce NC temp data
```{r NC temp data reduce}
head(NC_temp)

#fix date
NC_temp$Date <- ymd(NC_temp$date)

#I notice that we have many measurements for a single day, let's take the average for each day
NC_temp.r <- NC_temp %>% #take NC_temp data frame and then...
  group_by(Date) %>% #group by Date of measurement
  summarise(SST = mean(WTMP_avg)) #take the average value for each Date and put into column SST (sea surface temperature)

#also, add new region column
NC_temp.r$region <- "NC"
```


```{r NC population column type change}
NC_popcatchid$Date <- ymd(NC_popcatchid$Date)
```


```{r NJ population column type change }
NJ_popcatchid$Date <- ymd(NJ_popcatchid$Date)
```

```{r prepping measurement data for new jersey measurements}
head(NJfishmeasurements)

#get rid of ID column 1
NJfishmeasurements.r <- NJfishmeasurements[,-1] #before coma=rows after=columns - delete


#NJ_fishfilenames <- list.files("C:/Users/15512/Downloads/ij153-win-java8/ImageJ/FishlarvePICS") #may have to go into fileexplore and delete 204
#NJ_fishfilenames.r <- NJ_fishfilenames[1:441]
#save(NJ_fishfilenames.r, file = "NJ_fishfilenames.RData")
load("NJ_fishfilenames.RData")
nrow(NJfishmeasurements.r)
NJfishmeasurements.r$Code <- NJ_fishfilenames.r #would not work if the 441 didnot match up in measurements
view(NJfishmeasurements.r)

#Hi Justin! Previously, we were using gsub here, but I realized we could just use substr like we did for North Carolina, so it's simplified a bit

NJfishmeasurements.r$code2 <- substr(NJfishmeasurements.r$Code,1, 5) #only keep first 5 digits in string
NJfishmeasurements.r <- subset(NJfishmeasurements.r, select= -c(Code))
view(NJfishmeasurements.r)
NJfishmeasurements.r$Code <- NJfishmeasurements.r$code2 #change name back to Code to match

NJfishmeasurements.final <- NJfishmeasurements.r %>% 
  group_by(Code) %>% 
summarise(across(.cols=1:interorbitalwidth, .fns=~mean(.x, na.rm = TRUE), .names = "{.col}"))

  
```

Preparing measurement data for NC measurements
```{r prepping measurement data for North Carolina measurements}
head(NCfishmesurements)

#get rid of ID column 1
NCfishmeasurements.r <- NCfishmesurements[,-1] #before coma=rows after=columns - delete


#NC_fishfilenames <- list.files("C:/Users/15512/Downloads/ij153-win-java8/ImageJ/NC fish pics")
#NC_fishfilenames.r <- NC_fishfilenames[1:93]
#save(NC_fishfilenames.r, file = "NC_fishfilenames.RData")
load("NC_fishfilenames.RData")

nrow(NCfishmeasurements.r)
NCfishmeasurements.r$Code <- NC_fishfilenames.r #would not work if the 93 did not match up in measurements
NCfishmeasurements.r$Sample.ID <- substr(NCfishmeasurements.r$Code, 17,25)
NCfishmeasurements.r$Photo.ID <- substr(NCfishmeasurements.r$Code, 1,15)
NCfishmeasurements.r$Photo.ID2 <- substr(NCfishmeasurements.r$Code, 1,12)
NCfishmeasurements.r$Fish.ID <- paste(NCfishmeasurements.r$Photo.ID2,NCfishmeasurements.r$Sample.ID,sep = "_")


NCfishmeasurements.final <- NCfishmeasurements.r %>% 
  group_by(Fish.ID) %>% 
summarise(across(.cols=1:Eyedimatere, .fns=~mean(.x, na.rm = TRUE), .names = "{.col}"))
```


Bring in photo keys for North Carolina and for New Jersey
```{r bring in photo keys fix date}
#north carolina
NC_key_photoid <- read.csv("NC_key_photoid.csv")

#fix date
NC_key_photoid$Date <- mdy(NC_key_photoid$ï..Date.Collected)

#what are all of the column names
head(NC_key_photoid) #keep all

#add new column for region (NC versus NJ)
NC_key_photoid$region <- "NC"

#new jersey
NJ_key_photoid <- read.csv("NJ_2017_Photokey.csv")

#fix date
NJ_key_photoid$Date <- mdy(NJ_key_photoid$Tow.Date)

#what are the column names
head(NJ_key_photoid)

#we only want Code, Tow ID and Date
NJ_key_photoid.r <- NJ_key_photoid %>%
  select(Code,Date)

# we also want to add a new column for region (NC versus NJ)
NJ_key_photoid.r$region <- "NJ"

```

Combine all tables we need for NJ: temperature, photo key, and measurements
```{r combine NJ data tables}
#temperature: NJ_temp.r
#measurements: NJfishmeasurements.final
#photokey: NJ_key_photoid.r

#what are the column names again?
head(NJ_temp.r) 
#let's change SuTemp to SST
names(NJ_temp.r) <- c("Date","SST","region")

head(NJfishmeasurements.final)
NJfishmeasurements.final$Code <- as.numeric(NJfishmeasurements.final$Code)
head(NJ_key_photoid.r)

view(NJfishmeasurements.final)

```

Merge tables for NJ
```{r merge tables NJ}
#merge photo id with measurements
NJ_photoid_measurements <- inner_join(NJfishmeasurements.final,NJ_key_photoid.r, by = "Code")

#merge above with temp data

NJ_data_final <- inner_join(NJ_photoid_measurements, NJ_temp.r, by = c("Date","region"))

View(NJ_data_final)

```


Combine all tables we need for NC
```{r combine NC data tables}
#temperature: NC_temp.r
#measurements: NCfishmeasurements.final
#photokey: NC_key_photoid

#what are the column names again?
head(NC_temp.r) 
NC_temp.r$region <- "NC"
head(NCfishmeasurements.final)
NCfishmeasurements.final$region <- "NC"

head(NC_key_photoid)
```

Merge tables for NC
```{r merge tables NC}
#merge photo id with measurements
NC_photoid_measurements <- inner_join(NCfishmeasurements.final, NC_key_photoid, by = c("Fish.ID","region"))

#merge above with temp data

NC_data_final <- inner_join(NC_photoid_measurements, NC_temp.r, by = c("Date","region"))

colnames(NC_photoid_measurements)
colnames(NC_temp.r)

```

Match column names and order from two data frames
```{r}
colnames(NJ_data_final)
colnames(NC_data_final)
colnames(NC_data_final)[16] <- "Date.Collected"

```

NJ:
 [1] "Code"                "standardlength"      "totallength"         "bodydepthpectoral"  
 [5] "bodydepthanus"       "propulsivearea"      "bodyarea"            "musclearea"         
 [9] "eyediameter"         "caudalpeduncledepth" "caudalfinheight"     "caudalfinarea"      
[13] "bodywidth"           "interorbitalwidth"   "Date"                "region"             
[17] "SST" 

NC:
 [1] "Fish.ID"          "Standardlength"   "Totallength"      "Depthatpec"      
 [5] "Depthatanus"      "Pulsearea"        "Musclearea"       "Bodyarea"        
 [9] "Cauddepth"        "Caudheight"       "Caudarea"         "Bodywidth"       
[13] "Interoctdistance" "Eyedimatere"      "region"           "Date.Collected"  
[17] "Sample.ID"        "Photo.ID"         "Photo_reduced"    "Date"            
[21] "SST" 

First, match # and order of columns
```{r match # and order of NC columns}
NC_data_final.reorder <- NC_data_final[,c(1,2,3,4,5,6,8,7,14,9,10,11,12,13,20,15,21)]
colnames(NC_data_final.reorder)
```
Second, match names
```{r}
names(NC_data_final.reorder) <- names(NJ_data_final)
```


NJ | NC
Code | Fish.ID
standardlength | Standardlength
totallength | Totallength
bodydepthpectoral | Depthatpec
bodydepthanus | Depthatanus
propulsivearea | Pulsearea
bodyarea | Bodyarea
musclearea | Musclearea
eyediameter | Eyedimatere
caudalpeduncledepth | Cauddepth
caudalfinheight | Caudheight
caudalfinarea | Caudarea
bodywidth | Bodywidth
interorbitalwidth | Interoctdistance
Date | Date
region | region
SST | SST


Finally, we need to merge the NJ data table with the NC datatable

```{r final merge}
larval_data_final <- rbind(NC_data_final.reorder, NJ_data_final)

#save(larval_data_final, file = "larval_data_final.RData")

load("larval_data_final.RData")
Larvalpopulationdatafinal <- rbind(NJ_popcatchid, NC_popcatchid)
```

Fisher and Hogan 2007 Highlight some important ratios, so we'll calculate those as well. Specifically, we want to look at "Overall, we found that a predictive model incorporating total length (TL), the square of caudal peduncle depth factor (CPDF2) and aspect ratio (AR) can be used to accurately predict swimming performance of a wide range of fish families..."

Morphometric ratios
- Muscle ratio (MR) MA/BA
- Propulsive ratio (PR) PA/BA
- Fineness ratio (FR) TL/[(BW+BD)/2]
- Aspect ratio (AR) (CFD/CFA) ^(1/2)
- Caudal peduncle depth factor (CPDF) CPD/BD** note, was this  measured correctly?

```{r add ratio columns }
larval_data_final$muscle_ratio <- larval_data_final$musclearea/larval_data_final$bodyarea
larval_data_final$propulsive_ratio <- larval_data_final$propulsivearea/larval_data_final$bodyarea
larval_data_final$fineness_ratio <- larval_data_final$totallength/((larval_data_final$bodywidth + larval_data_final$bodydepthpectoral)/2)
larval_data_final$aspect_ratio <- larval_data_final$caudalpeduncledepth/larval_data_final$caudalfinarea^0.5
larval_data_final$caudal_peduncle_dfactor <- larval_data_final$caudalpeduncledepth/larval_data_final$bodydepthpectoral


```

---------

##Data Analysis for location, date, temperature, number of fish, and morphology

Question: Is there a correlation between aspect ratio and SST? 
```{r}
ggplot(larval_data_final) + 
  geom_point(aes(x = SST, y = aspect_ratio, color = region)) + 
  geom_smooth(aes(x = SST, y = aspect_ratio, color = region), method= "lm") +
  theme_dark()
  
ggplot(larval_data_final) + 
  geom_point(aes(x = SST, y = aspect_ratio, color = region)) + 
  geom_smooth(aes(x = SST, y = aspect_ratio, color = region), method= "lm") +
  theme_get()
```

Question: Is there relation between total lengeth and date? 
```{r}
ggplot(larval_data_final) + 
  geom_point(aes(x = SST, y = totallength, color = region)) + 
  geom_smooth(aes(x = SST, y = totallength, color = region), method= "lm") +
   xlim(10,30)
  theme_dark()
```

Is there a relationship between fitnessratio and date? 
```{r}
ggplot(larval_data_final) + 
  geom_point(aes(x = SST, y = fineness_ratio, color = region)) + 
  geom_smooth(aes(x = SST, y = fineness_ratio, color = region), method= "lm") +
    xlim(10,30)
  theme_dark()
```

Question? What is distrubituion of standards lengths in NJ? 

Graph
```{r}
ggplot(NJ_data_final) + 
  geom_histogram(aes(x = standardlength)) +
  theme_classic()
```
Question? What is distrubituion of standards lengths in NC? 

Graph
```{r}
ggplot(NC_data_final.reorder) + 
  geom_histogram(aes(x = standardlength)) +
  theme_classic()
```
Plotting together!
```{r}
ggplot(larval_data_final) + 
  geom_histogram(aes(x = standardlength)) +
  facet_wrap(~region) + #one graph for each region
  theme_classic()
```
```

Statistics

-----------
Question? How does caudal height change over time? 

Graph
```{r}
ggplot(larval_data_final) + 
  geom_point(aes(x = Date, y = caudalfinheight, color = region)) + 
  geom_smooth(aes(x = Date, y = caudalfinheight, color = region), method= "lm") +
  theme_classic()
```


Statistics

Question: How does caudal height correlate with temperature?

```{r}
ggplot(larval_data_final) + 
  geom_point(aes(x = SST, y = caudalfinheight, color = region)) + 
  geom_smooth(aes(x = SST, y = caudalfinheight, color = region), method= "lm") +
  xlim(10,30) + #sets min and max
  theme_classic()
```

Question: Is the Muslce ratio in NJ and NC consistant over time? 
```{r}
ggplot(larval_data_final) + 
  geom_point(aes(x = Date, y = muscle_ratio, color = region)) + 
  geom_smooth(aes(x = Date, y = muscle_ratio, color = region), method= "lm") +
  theme_classic()
  
  ggplot(larval_data_final) +
   geom_boxplot(aes(x = region, y = muscle_ratio)) +
  theme_classic()
  
  larval_data_final_NJ <- larval_data_final %>% filter(region == "NJ")
  larval_data_final_NC <- larval_data_final %>% filter(region == "NC")

  
  t.test(larval_data_final_NJ$muscle_ratio, larval_data_final_NC$muscle_ratio)

view(larval_data_final)
```


---------

Question: How does caudal height correlate with temperature?

```{r}
ggplot(larval_data_final) + 
  geom_point(aes(x = SST, y = caudalfinheight, color = region)) + 
  geom_smooth(aes(x = SST, y = caudalfinheight, color = region), method= "lm") +
  xlim(10,30) + #sets min and max
  theme_classic()
```
Is there a relationship between population size and sst?
```{r}
ggplot(larval_data_final) + 
  geom_point(aes(x = n, y = SST, color = region)) + 
  geom_smooth(aes(x = n , y = SST, color = region), method= "lm") +
  theme_classic()
```


Question: Is the Caudalpadundactor in NJ and NC consistant over time? 
```{r}
ggplot(larval_data_final) + 
  geom_point(aes(x = Date, y = caudal_peduncle_dfactor, color = region)) + 
  geom_smooth(aes(x = Date, y = caudal_peduncle_dfactor, color = region), method= "lm") +
  theme_dark()
  
  ggplot(larval_data_final) +
   geom_boxplot(aes(x = region, y = caudal_peduncle_dfactor)) +
  theme_classic()
  
  t.test(larval_data_final_NJ$caudal_peduncle_dfactor, larval_data_final_NC$caudal_peduncle_dfactor)

view(larval_data_final)
```

Question: Is the Aspectratio in NJ and NC consistant over time? 
```{r}
ggplot(larval_data_final) + 
  geom_point(aes(x = Date, y = aspect_ratio, color = region)) + 
  geom_smooth(aes(x = Date, y = aspect_ratio, color = region), method= "lm") +
  theme_bw()
  
  ggplot(larval_data_final) +
   geom_boxplot(aes(x = region, y = aspect_ratio)) +
  theme_bw()
  
  t.test(larval_data_final_NJ$aspect_ratio, larval_data_final_NC$aspect_ratio)

view(larval_data_final)
```


Question: Is the Total Lengeth in NJ and NC consistant over time? 
```{r}
ggplot(larval_data_final) + 
  geom_point(aes(x = Date, y = totallength, color = region)) + 
  geom_smooth(aes(x = Date, y = totallength, color = region), method= "lm") +
  theme_bw()
  
  ggplot(larval_data_final) +
   geom_boxplot(aes(x = region, y = totallength)) +
  theme_bw()
  
  t.test(larval_data_final_NJ$totallength, larval_data_final_NC$totallength)

Question: Is there a relationship between population size and date?


```{r}
ggplot(Larvalpopulationdatafinal) +
geom_point(aes(x = Date, y = log(n) ))+
facet_wrap(~region) +
theme_()

ggplot(Larvalpopulationdatafinal) +
geom_point(aes(x = Date, y = log(n), color=region ))+
theme_dark()
```

Question: Is there a relationship between date and temperature?
```{r}
  
  ggplot(larval_data_final) +
   geom_boxplot(aes(x = region, y = SST)) +
  ylim(0, 30)+
  theme_bw()
  
  t.test(larval_data_final_NJ$SST, larval_data_final_NC$SST)

  ggplot(larval_data_final) +
geom_smooth(aes(x = region, y = SST , color=region ))+
facet_wrap(~region) +
ylim(0, 30)+
theme_bw()

ggplot(Larvalpopulationdatafinal) +
geom_point(aes(x = Date, y = log(n), color=region ))+
theme_bw()
```
---------
Use ggplot() to graph, this [CheatSheet](https://www.maths.usyd.edu.au/u/UG/SM/STAT3022/r/current/Misc/data-visualization-2.1.pdf) will be helpful!



Some statistical tests that will be helpful:

Are these two groups different?
- # independent 2-group t-test
- t.test(y1,y2)

To visualize different variables it may be helpful to use:
- scatter plots (i.e.Date versus morphology)
- PCA (does fish morphology group by region)
- box plots (i.e.Region versus morphology)